\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2018-02-19 redaction/dhp_july2017.tex
Levinshon The Box

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se} (\citet{daudin_logistique_2003, daudin_transactions_2005}). In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs would represent a 21\% markup over production costs. Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) suggests a non-negligible impact of transport costs on trade flows. If much of trade policy barriers have been removed over the second half of the 20th century, this suggests that the transport costs component remains large and deserves attention. International transport costs accordingly stand at the heart of the paper.\smallskip

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others). Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms. However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much. In this paper, we investigate the robustness of this result. As major contribution of the paper, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs. Rather, transport costs ``per se'' did decline, in both transport modes, and this is constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.\medskip

>>>>

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se} (\citet{daudin_logistique_2003, daudin_transactions_2005}). In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs would represent a 21\% markup over production costs. Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) indicates that the impact of transport costs on trade flows is not negligible. If much of trade policy barriers have been removed over the second half of the 20th century, this suggests that the transport costs component remains large and deserves attention. International transport costs accordingly stand at the heart of the paper.\smallskip

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others). Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). %According to \cite{Glaeser04}, over the twentieth century, the cost of moving goods inland have declined by over 90\% in real terms.
The container was a major transport-cost reducing innovation (see \citealp{levinson2016box}). However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much. In this paper, we investigate the robustness of this result. As major contribution of the paper, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs. Rather, transport costs ``per se'' did decline, in both transport modes, and this is constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.\medskip
